A Gaussian process model of human electrocorticographic data

### 2020-02-20 paper/boneyard.tex
finished draft of results section

>>>>

as $\delta$, $\theta$, and $\alpha$, have been implicated in processing modes
which span large swaths of cortical regions and are thought to serve the purpose
of integration across many cortical sites ~\citep{Knya12}. Higher frequency
oscillations such as $\beta$ and $\gamma$, are more limited to topographic areas
~\citep{Nune95}, and are modulated by slow oscillations hierarchically
~\citep{EngeEtal10}. Using the frequency-specific reconstruction accuracies, we
followed the process outlined in Fig.~\ref{fig:informap}C, to obtain the most
informative locations for each frequency.  When we examined the intersection
between the top 10\% most informative voxels across the two datasets for each
bandpassed frequency, we found a somewhat consistent hierarchical structure (see
Fig.~\ref{fig:networks}E).

; $\theta$ activity
in frontal regions have been implicated in memory
consolidation~\citep{NishEtal09} and integration~\citep{BackEtal16}; parietal
and occipital $\alpha$ activity has been implicated in attention
modulation~\citep[e.g., see][]{Klim12}; $\beta$ activity in motor cortex have
been implicated in motor control and processing~\citep{BrovEtal04, KoelEtal08,
RubiEtal06}; and occipital $\gamma$ activity has been implicated in visual
attention and visual processing~\citep{ZhouEtal16, TallEtal05, FrieEtal07}.

%next: neurosynth

%%%%STOPPED HERE

We then wondered which networks were involved in each of these
frequency-specific intersecting regions.  To better understand the networks
involved we colored the regions (Fig.~\ref{fig:networks}A) based on the network
parcellation ~\citep[using $k$=7;][]{YeoEtal11}} (Fig.~\ref{fig:networks}D) and
calculated the proportion of each of those networks (Fig.~\ref{fig:networks}B).
Consistent with the idea outlined above, we found networks most involved in the
lowest frequency band, $\delta$, were networks involved integration and global
processing (specifically the dorsal attention ~\citep{MarzEtal13}, default mode,
and frontoparietal networks).  Although these networks are also seen in the
$\theta$ intersection, they are less proportionally to the visual network.
Alpha power is consistently modulated in EEG studies of attention and
visuo-motor ~\citep{MantEtal07, MarzEtal13}, and we similarly found a high
proportion of visual, somatomotor, and dorsal attention networks within this
band.  We found $\beta$ intersections to predominantly consist of somatomotor
network, in line with~\citep{RittEtal09, VaneEtal11}.  Also accordant with
heavily cited literature~\citep{FrieEtal07}, the network most predominant in the
$\gamma$ intersection was the visual network.  Lastly, broadband power was found
predominantly in frontoparietal and default mode networks, consistent with
~\citep{OssaEtal11}.

### 2020-02-20 paper/main.tex
backing up progress
finished draft of results section

We also wondered whether the map of information scores might vary as a function of the spectral
components of the activity patterns under consideration.  For example, low frequency $\delta$ activity in the thalamus has been implicated in
supporting connectivity with cortex~\citep[e.g.,][]{Jone09}; $\theta$ activity
in frontal regions have been implicated in memory consolidation~\citep{NishEtal09} and integration~\citep{BackEtal16}; parietal and occipital $\alpha activity has been
implicated in attention modulation~\citep[e.g., see][]{Klim12}; $\beta$ activity in motor cortex have been implicated in motor control and processing~\citep{BrovEtal04, KoelEtal08, RubiEtal06};
and occipital $\gamma$ activity has been implicated in visual attention and visual processing~\citep{ZhouEtal16, TallEtal05, FrieEtal07}.
We computed analogous maps of information scores for each individual frequency band.
Across Datasets 1 and 2 (with the exception of $\alpha$ band activity), we observed reliable positive correlations
between the voxel-by-voxel maps of information scores ($\delta$: $r = 0.09, p < 10^{-57}$;
$\theta$: $r = 0.24, p < 10^{-60}$; $\alpha$: $r = -0.03, p < 0.001$,
$\beta$: $r = 0.02, p - 0.0011$; $\gamma_L$: $r = 0.1, p < 10^{-67}$;
$\gamma_H$: $r = 0.03, p < 10^{-7}$; broadband: $r = 0.21, p < 10^{-297}).

@@

A growing body of neuroscientific research is concerned with characterizing the \textit{parcellations}
of anatomical and functional brain networks~\citep[for review see][]{ZaleEtal10, ArslEtal18}.
The dominant approaches entail obtaining a full-brain connectivity matrix using either diffusion tensor imaging to identify the brain's network of
white matter connections, or functional connectivity (typically applied to resting state data) to
correlate the patterns of activity exhibited by different brain structures.  One can then apply graph theoretic approaches
to assign each brain structure (typically a single fMRI voxel) to one or more networks~\citep[for review see][]{BullSpor09}.
The result is a set of distinct (or partially overlapping) brain ``networks'' that may be
further examined to elucidate their potential functional role.  We overlaid a well-cited seven-network
parcellation map identified by \cite{YeoEtal11} onto the maps of brain locations that were most informative
about each neural pattern.  For each of these information maps, we computed the proportion of voxels in the most informative
brain regions that belonged to each of the seven networks identified by \cite{YeoEtal11}.
We found that the regions we identified as being most informative about different neural patterns varied
markedly with respect to which functional networks they belonged to (Fig.~\ref{fig:networks}A, B).  This is consistent with
the notion that there is no ``universal'' brain region that consistently reflects activity in the rest of the brain.  Rather,
each region's activity patterns appear to be characterized by different spectral profiles, and the ability to
infer activity patterns at a particular frequency range depends on the structural and functional
connectome specific to that frequency range.

%next: neurosynth

%%%%STOPPED HERE

as $\delta$, $\theta$, and $\alpha$, have
been implicated in processing modes which span large swaths of
cortical regions and are thought to serve the purpose of integration
across many cortical sites ~\citep{Knya12}. Higher frequency
oscillations such as $\beta$ and $\gamma$, are more limited to topographic
areas ~\citep{Nune95}, and are modulated by slow oscillations
hierarchically ~\citep{EngeEtal10}. Using the frequency-specific reconstruction accuracies, we
followed the process outlined in Fig.~\ref{fig:informap}C, to obtain
the most informative locations for each frequency.  When we examined the intersection
between the top 10\% most informative voxels across the two datasets
for each bandpassed frequency, we found a somewhat consistent hierarchical
structure (see Fig.~\ref{fig:networks}E).

We then wondered which
networks were involved in each of
these frequency-specific intersecting regions.  To better understand the networks involved
we colored the regions (Fig.~\ref{fig:networks}A) based on the
network parcellation ~\citep[using $k$=7;][]{YeoEtal11}} (Fig.~\ref{fig:networks}D) and calculated the proportion of each
of those networks (Fig.~\ref{fig:networks}B).  Consistent with the idea outlined
above, we found networks most involved in the lowest frequency band,
$\delta$, were networks involved integration and global processing
(specifically the dorsal attention ~\citep{MarzEtal13}, default mode,
and frontoparietal networks).  Although these networks are also seen
in the $\theta$ intersection, they are less proportionally to the
visual network.  Alpha power is consistently modulated in EEG studies
of attention and visuo-motor ~\citep{MantEtal07, MarzEtal13}, and we
similarly found a high proportion of visual, somatomotor, and dorsal attention networks
within this band.  We found $\beta$ intersections to predominantly
consist of somatomotor network, in line with~\citep{RittEtal09, VaneEtal11}.  Also accordant with heavily cited
literature~\citep{FrieEtal07}, the network most predominant in the
$\gamma$ intersection was the visual network.  Lastly, broadband power
was found predominantly in frontoparietal and default mode networks,
consistent with ~\citep{OssaEtal11}.

>>>>

We also wondered whether the map of information scores might vary as a function
of the spectral components of the activity patterns under consideration.  For
example, low frequency $\delta$ activity in the thalamus has been implicated in
supporting connectivity with cortex~\citep[e.g.,][]{Jone09}. We computed
analogous maps of information scores for each individual frequency band. Across
Datasets 1 and 2 (with the exception of $\alpha$ band activity), we observed
reliable positive correlations between the voxel-by-voxel maps of information
scores ($\delta$: $r = 0.09, p < 10^{-57}$; $\theta$: $r = 0.24, p < 10^{-60}$;
$\alpha$: $r = -0.03, p < 0.001$, $\beta$: $r = 0.02, p - 0.0011$; $\gamma_L$:
$r = 0.1, p < 10^{-67}$; $\gamma_H$: $r = 0.03, p < 10^{-7}$; broadband: $r =
0.21, p < 10^{-297})$.

@@

A growing body of neuroscientific research is concerned with characterizing the
\textit{parcellations} of anatomical and functional brain networks~\citep[for
review see][]{ZaleEtal10, ArslEtal18}. The dominant approaches entail obtaining
a full-brain connectivity matrix using either diffusion tensor imaging to
identify the brain's network of white matter connections, or functional
connectivity (typically applied to resting state data) to correlate the patterns
of activity exhibited by different brain structures.  One can then apply graph
theoretic approaches to assign each brain structure (typically a single fMRI
voxel) to one or more networks~\citep[for review see][]{BullSpor09}. The result
is a set of distinct (or partially overlapping) brain ``networks'' that may be
further examined to elucidate their potential functional role.  We overlaid a
well-cited seven-network parcellation map identified by \cite{YeoEtal11} onto
the maps of brain locations that were most informative about each neural
pattern.  For each of these information maps, we computed the proportion of
voxels in the most informative brain regions that belonged to each of the seven
networks identified by \cite{YeoEtal11}; Figure~\ref{fig:networks}D. We found
that the regions we identified as being most informative about different neural
patterns varied markedly with respect to which functional networks they belonged
to (Fig.~\ref{fig:networks}A, B).

The frequency-specific information score maps we observed were consistent with
the notion that there is no ``universal'' brain region that  reflects all
activity patterns throughout the rest of the brain.  Rather, each region's
activity patterns appear to be characterized by different spectral profiles, and
the ability to infer full-brain activity patterns at a particular frequency
range depends on the structural and functional connectome specific to that
frequency range (Fig.~\ref{fig:networks}E).  We wondered how the maps we found
might fit in with prior work.  To this end, in addition to examining the
anatomical profiles of each map, we used Neurosynth~\citep{RubiEtal17} to
identify (using meta analyses of the neuroimaging literature) the top five most
common terms associated with each frequency-specific map
(Fig.~\ref{fig:networks}C).  We found that $\delta$ patterns across the brain
were best predicted by regions of ventromedial prefrontal cortex, striatum, and
thalamus (yellow).  These regions are also implicated in modulating $\delta$
oscillations during sleep, and are heavily interconnected with
cortex~\citep[e.g.,][]{AmziSter98}.  The brain areas most predictive of
full-brain $\theta$ patterns were occipital and parietal regions traditionally
linked with visual processing and visual attention (light green). Prior work has
implicated $\theta$ oscillations in these areas in periodic sampling of visual
attention~\citep[e.g.,][]{BuscVanR10}.  We found that full-brain $\alpha$
patterns were best predicted by motor areas (dark green), which also exhibit
$\alpha$ band changes during voluntary movements~\citep[e.g.,][]{JurkEtal06}.
Striatum and thalamus (teal) were most informative about full-brain $\beta$
patterns.  Prior work has implicated striatal $\beta$ activity in sensory and
motor processing~\citep{FeinEtal15} and thalamic $\beta$ activity has been
implicated in modulating widespread $\beta$ patterns across
neocortex~\citep{SherEtal16}. Somatosensory areas (dark blue) were most
informative about full-brain $\gamma_L$ patterns. Prior work has implicated
somatosensory $\gamma_L$ in somatosensory processing and motor
planning~\citep{AyaEtal03}.  Occipital cortex (purple) was most informative
about full-brain $\gamma_H$ patterns.  Occipital $\gamma_H$ has also been linked
with visual processing and reading~\citep{WuEtal11} and the transmission of
visual representations from low-order to higher-order visual
areas~\citep{MatsEtal13}.  Full-brain broadband patterns were best predicted by
inferior parietal cortex precuneus (maroon).  Functional neuroimaging BOLD
responses~\citep{SimoEtal16} and broadband ECoG patterns~\citep{HoneEtal12} in
these default-mode hubs have been implicated in processing context-dependent
representations that unfold over long timescales.

Taken together, the frequency-specific information maps suggest potential new
interpretation of many of the above previously reported findings.  Prior work
has largely treated region-specific narrowband and broadband activity as an
indictor that activity at those frequency ranges reflect that the given region
is representing or supporting a particular function.  Our work suggests the
alternative interpretation that when we observe a particular neural pattern in a
particular brain region, it may instead reflect how that region is is transmitting
information to the rest of the brain via patterns at the given frequency range.
